---
title: "Urban Heat Island Detection & Impact Analysis - Dallas (2022)"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

## Dataset Description

Climatological, hourly time-series data has been collected from several weather stations run by the NWS, FAA, and DOD, including airports and other significant areas, is compiled into 3 Local Climatological Datasets (LCD) for Arlington, Dallas, and Denton for 2022.

The dataset also offers daily summaries that describe:

| Class                | Description       |
|----------------------|-------------------|
| Wind Conditions      | (Add description) |
| Precipitation Levels | (Add description) |
| Degree Days          | (Add description) |
| Temperature Extremes | (Add description) |

Additional data is included in the tabulated monthly summaries: maximum, minimum, and average temperatures; deviations from normal; dew point temperatures; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; relative humidity; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth.

#### Dallas 2022 Dataset

```{r}
#| label: dallas-dataset
#| echo: false
#| warning: false

dallas_data <- read_csv(here('./data/dallas-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(dallas_data)
num_columns <- ncol(dallas_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")


dallas_data |>
  head() |>
  formattable()
```

#### **Arlington 2022 Dataset**

```{r}
#| label: arlington-dataset
#| echo: false
#| warning: false

arlington_data <- read_csv(here('./data/arlington-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(arlington_data)
num_columns <- ncol(arlington_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

arlington_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

### **Denton 2022 Dataset**

```{r}
#| label: denton-dataset
#| echo: false
#| warning: false

denton_data <- read_csv(here('./data/denton-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(denton_data)
num_columns <- ncol(denton_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

denton_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

## Questions

1.  Can we identify specific anomalies that are characteristic of urban heat islands, and how do these anomalies correlate with urban infrastructure and population density?

2.  How do urban heat islands vary in intensity across different metropolitan/rural areas, and what are the temporal patterns of these variations within a given year?

## Analysis plan

Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 12/6/2023)
