---
title: "Urban Heat Island Detection & Impact Analysis - Dallas (2022)"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

## Dataset

*A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.*

*Make sure to load the data and use inline code for some of this information.*

[**Byte Busters' Draft:**]{.underline}

The Byte Busters will be using data from the National Centers for Environmental Information

(<https://www.ncei.noaa.gov/>)...

[Udit & Ajinkya:]{.underline}

```{r}
#| label: load-dallas-dataset
#| echo: false
#| warning: false

dallas_data <- read_csv(here('./data/dallas-2022.csv'))

dallas_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

[Riyanshi]{.underline}:

-   *Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.*

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 13/6/2023)
