---
title: "Urban Heat Island Detection & Impact Analysis - Dallas (2022)"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

## Dataset

Data Description: The climatological data collected from several weather stations run by the NWS, FAA, and DOD, including airports and other significant areas, is compiled into 3 Local Climatological Datasets (LCD) for Arlington, Dallas, and Denton for the year 2022.

The dataset also offers daily summaries that describe wind conditions, precipitation levels, degree days, and temperature extremes.Data like maximum, minimum, and average temperatures; deviations from normal; dew point temperatures; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; relative humidity; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth are all included in the tabulated monthly summaries.

### **Dallas 2022 Dataset**

```{r}
#| label: dallas-dataset
#| echo: false
#| warning: false

dallas_data <- read_csv(here('./data/dallas-2022.csv'))

dallas_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
dim(dallas_data)

dallas_dimension <- data.frame(dallas_data)
dallas_dimension

num_rows <- nrow(dallas_data)
num_rows
num_columns <- ncol(dallas_data)
num_columns
```

### **Arlington 2022 Dataset**

```{r}
#| label: arlington-dataset
#| echo: false
#| warning: false

arlington_data <- read_csv(here('./data/arlington-2022.csv'))

arlington_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

### **Denton 2022 Dataset**

```{r}
#| label: denton-dataset
#| echo: false
#| warning: false

denton_data <- read_csv(here('./data/denton-2022.csv'))

denton_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

[Riyanshi]{.underline}:

-   *Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.*

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 13/6/2023)
