---
title: "Urban Heat Island Detection & Impact Analysis - Dallas (2022)"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

##  Project Objective**:**

-   The primary objective of this project is to leverage data mining and machine learning algorithms to detect and characterize urban heat islands (UHIs) across a diverse range of metropolitan landscapes. By analyzing weather pattern anomalies and temperature distributions, the project aims to gain a deeper understanding of the microclimatic effects of urbanization.

#### **Scope and Focus:**

-   This project will focus on assessing UHIs in various urban environments, particularly comparing the UHI intensity in the following settings:

#### **Dallas (Significant City):**

-   Analyzing the UHI in a major metropolitan area with high population density and urbanization. Factors such as pollution, land use, and climate will be considered to determine the UHI's intensity.

#### **Arlington (Suburban Town):**

-   Evaluating the UHI in a suburban setting, with a moderate population density and less urban development than a significant city. This comparison will help us understand how UHIs vary in less densely populated areas.

#### **Denton (Rural City):**

-   Examining the UHI in a rural city, where factors like lower population density, less pollution, and different land use patterns play a role in UHI intensity.

#### Dataset Description:

The high-level goal of this project is to detect, analyze, and understand Urban Heat Islands (UHIs) by leveraging data mining and machine learning to understand the microclimatic effects of urbanization in various metropolitan landscapes.

Climatological, hourly time-series data has been collected from several weather stations run by the National Weather Service (NWS), Federal Aviation Administration (FAA), and Department of Defense (DOD), including airports and other significant areas, is compiled into 3 Local Climatological Datasets (LCD) for Arlington, Dallas, and Denton for 2022.

The dataset also offers daily summaries that describe:

| Columns              | Description       |
|----------------------|-------------------|
| Wind Conditions      | (Add description) |
| Precipitation Levels | (Add description) |
| Degree Days          | (Add description) |
|                      |                   |

Additional data is included in the tabulated monthly summaries: maximum, minimum, and average temperatures; deviations from normal; dew point temperatures; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; relative humidity; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth.

#### Dallas 2022 Dataset

```{r}
#| label: dallas-dataset
#| echo: false
#| warning: false

dallas_data <- read_csv(here('./data/dallas-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(dallas_data)
num_columns <- ncol(dallas_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")


dallas_data |>
  head() |>
  formattable()
```

#### **Arlington 2022 Dataset**

```{r}
#| label: arlington-dataset
#| echo: false
#| warning: false

arlington_data <- read_csv(here('./data/arlington-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(arlington_data)
num_columns <- ncol(arlington_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

arlington_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

### **Denton 2022 Dataset**

```{r}
#| label: denton-dataset
#| echo: false
#| warning: false

denton_data <- read_csv(here('./data/denton-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(denton_data)
num_columns <- ncol(denton_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

denton_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

## Questions

1.  How can we pinpoint distinct anomalies that are typical of urban heat islands, and what is the extent of their correlation with urban infrastructure and population density?

2.  What methods can be employed to assess variations in the intensity of urban heat islands across diverse metropolitan and rural regions, and how can we characterize the temporal patterns of these variations throughout a specific year?

## Analysis plan

Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 12/6/2023)
