---
title: "Urban Heat Island Detection & Impact Analysis"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

## Project Objective**:**

The primary objective of this project is to leverage data mining and machine learning algorithms to detect and characterize Urban Heat Islands (UHIs) across three different areas in the Dallas region. By analyzing weather pattern anomalies and temperature distributions, the project aims to gain a deeper understanding of the microclimatic effects of urbanization.

#### **Scope and Focus:**

This project will focus on assessing UHIs in various urban environments, particularly comparing the UHI intensity in the following settings:

#### **Dallas (Significant City):**

Analyzing the UHI in a major metropolitan area with high population density and urbanization. Factors such as pollution, land use, and climate will be considered to determine the UHI's intensity.

#### **Arlington (Suburban Town):**

Evaluating the UHI in a suburban setting, with a moderate population density and less urban development than a significant city. This comparison will help us understand how UHIs vary in less densely populated areas.

#### **Denton (Rural City):**

Examining the UHI in a rural city, where factors like lower population density, less pollution, and different land use patterns play a role in UHI intensity.

#### Dataset Description:

The high-level goal of this project is to detect, analyze, and understand Urban Heat Islands (UHIs) by leveraging data mining and machine learning to understand the microclimatic effects of urbanization in various metropolitan landscapes.

Climatological, hourly time-series data has been collected from several weather stations run by the National Weather Service (NWS), Federal Aviation Administration (FAA), and Department of Defense (DOD), including airports and other significant areas, is compiled into 3 Local Climatological Datasets (LCD) for Arlington, Dallas, and Denton for 2022.

The dataset also offers daily summaries that describe:

| Columns                           | Description       |
|-----------------------------------|-------------------|
| Wind Gust Speed                   | Wind gusts occurring during time of observation. Given in miles per hour (mph)|
| Dry Bulb Temperature              | The measure of air temperature (Fahrenheit) |
| Precipitation Levels              | Total Liquid Content (TLC) Water equivalent amount of precipitation for the day |
| Dew Point Temperature             | (Add description) |
| Pressure - Avg SL                 | Daily average sea level pressure |
| Rel Hum                           |This is the relative humidity given to the nearest whole percentage|

Additional data is included in the tabulated monthly summaries: maximum, minimum, and average temperatures; deviations from normal; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth.

## Question

How do urban heat islands vary in intensity across different metropolitan/rural cities in a particular region, and what are the temporal patterns of these variations within a given year?

```{r}
# Q1, Part 1: k-Means Clustering, SVM (Support Vector Machines), Isolation Forest and Local Outlier Factor (LOF), Z-scores. First 3 are models.

# Q1, Part 2: Time-series analysis (data viz & dashboards; heat maps).
```

## Analysis Plan

Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 12/6/2023)

## Organization of the Project Repository

Below are the key folders within the Project repository:

-   **data/**: This directory is designated for essential data files required for the project, including input files.

-   **images/**: This folder serves as the storage location for project-related visual representations and plots.

-   **presentation_files/**: This folder is specifically reserved for materials relevant to project presentations.

-   **\_extra/**: We will brainstorm and store analysis-related content that doesn't directly affect the project workflow.

-   **\_freeze/**: This folder stores the generated files during the build process, preserving the website's specific state at a given moment.

-   **\_site/**: This is where we will generate our static website files after rendering our Quarto documents.

-   **.github/**: This is where we will access and manage GitHub templates and workflow configurations.
