```{r}

#| label: load-pkgs
#| include: false
#| message: false
#| warning: false

#install.packages("pacman")

pacman::p_load(dlookr,
               tidyverse,
               here,
               formattable,
               knitr,
               dplyr)
```

+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
| Week                                      | Weekly Tasks                                                               | Team Member Assignments     | Group Deliverables                                                      |
+:=========================================:+:==========================================================================:+:===========================:+:=======================================================================:+
| [**Week 1 (11/8 - 11/15)**]{.underline}\  | Setting up environment, data blending                                      | *Dhanya, Udit*              | Cleaned dataset ready for exploratory analysis                          |
| (Data Cleaning & Data Pre-processing)     |                                                                            |                             |                                                                         |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Identify missing values and perform imputation to deal with missing values | *Riyanshi, Ajinkya*         | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Data type conversions                                                      | *Kendall*                   | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
| [**Week 2 (11/15 - 11/22)**]{.underline}\ | Visualization of key metrics                                               | *Udit, Dhanya*              | A thorough analysis of data insights and prepared features for modeling |
| (Exploratory Data Analysis)               |                                                                            |                             |                                                                         |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Feature engineering and selection                                          | *Riyanshi*                  | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Data normalization and standardization                                     | *Kendall, Ajinkya*          | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
| [**Week 3 (11/22 - 11/29)**]{.underline}\ | Implement anomaly detection algorithms                                     | *Riyanshi, Kendall, Dhanya* | Anomaly detection results and cleaned dataset                           |
| (Anomaly Detection)                       |                                                                            |                             |                                                                         |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Validate and analyze anomalies                                             | *Udit, Ajinkya*             | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
| [**Week 4 (11/29 - 12/6)**]{.underline}\  | Build classification models                                                | *Dhanya, Kendall*           | Working models and performance evaluation report                        |
| (Model Development & Evaluation)          |                                                                            |                             |                                                                         |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Model optimization & cross-validate models                                 | *Riyanshi, Udit*            | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Evaluation metrics and performance analysis                                | *Ajinkya*                   | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
| [**Week 5 (12/6 - 12/6)**]{.underline}\   | Compile project documentation                                              | *Kendall, Dhanya*           | Complete project documentation and ready-to-deliver presentation        |
| (Documentation & Presentation)            |                                                                            |                             |                                                                         |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+
|                                           | Prepare final presentation                                                 | *Udit, Riyanshi, Ajinkya*   | \-                                                                      |
+-------------------------------------------+----------------------------------------------------------------------------+-----------------------------+-------------------------------------------------------------------------+

Make sure to load the data and use inline code for some of this information.

Byte Busters' Draft:

The Byte Busters will be using data from the National Centers for Environmental Information

(https://www.ncei.noaa.gov/)...

Udit & Ajinkya:

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

[Riyanshi]{.underline}:

-   *Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.*

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 13/6/2023)

\-\-\-\--

This is what Anjik sent to me on WhatsApp:

data1_rows, data1_cols = dallas_data.shape\
data2_rows, data2_cols = arlington_data.shape\
data3_rows, data3_cols = denton_data.shape

```{r}
#Dallas Dataset: 

dim(dallas_data)

dallas_dimension <- data.frame(dallas_data)
dallas_dimension
```

Anjinka & Udit

wind conditions, precipitation levels, degree days, and temperature extremes. Additional data is included in the tabulated monthly summaries: maximum, minimum, and average temperatures; deviations from normal; dew point temperatures; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; relative humidity; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth.

1.  What methods can be employed to assess variations in the intensity of urban heat islands across diverse metropolitan and rural regions, and how can we characterize the temporal patterns of these variations throughout a specific year?

<!-- -->

1.  How can we pinpoint distinct anomalies that are typical of urban heat islands, and what is the extent of their correlation with urban infrastructure and population density?

#### Dallas 2022 Dataset

`{r} #| label: dallas-dataset #| echo: false #| warning: false  dallas_data <- read_csv(here('./data/dallas-2022.csv'))  # Dimensions of dataset: num_rows <- nrow(dallas_data) num_columns <- ncol(dallas_data)  cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")   dallas_data |>   head() |>   formattable()}`

#### **Arlington 2022 Dataset**

```{r}
#| label: arlington-dataset 
#| echo: false 
#| warning: false  

arlington_data <- read_csv(here('./data/arlington-2022.csv'))  

# Dimensions of dataset: 
num_rows <- nrow(arlington_data) 
num_columns <- ncol(arlington_data)  
cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

arlington_data |> 
  head() |>   
  formattable()
```

### **Denton 2022 Dataset**

```{r}
#| label: denton-dataset
#| echo: false
#| warning: false

denton_data <- read_csv(here('./data/denton-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(denton_data)
num_columns <- ncol(denton_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

denton_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

-   **data/:** Used for storing any necessary data files for the project, such as input files.

-   **images/:** Used for storing plots of our project.

-   **presentation_files/:** Folder for having presentation related files.

-   **\_extra/:** Used to brainstorm our analysis which won't impact our project workflow.

-   **\_freeze/:** This folder is used to store the generated files during the build process. These files represent the frozen state of the website at a specific point in time.

-   **\_site/:** Folder used to store the generated static website files after the site generator processes the quarto document.

-   **.github/:** Folder for storing github templates and workflow.

+-----------------------+---------------------------------------------------------------------------------+
| Columns               | Description                                                                     |
+=======================+=================================================================================+
| Wind Gust Speed       | Wind gusts occurring during time of observation. Given in miles per hour (mph)  |
+-----------------------+---------------------------------------------------------------------------------+
| Dry Bulb Temperature  | The measure of air temperature (Fahrenheit)                                     |
+-----------------------+---------------------------------------------------------------------------------+
| Precipitation Levels  | Total Liquid Content (TLC) Water equivalent amount of precipitation for the day |
+-----------------------+---------------------------------------------------------------------------------+
| Dew Point Temperature | (Add description)                                                               |
+-----------------------+---------------------------------------------------------------------------------+
| Pressure - Avg SL     | Daily average sea level pressure                                                |
+-----------------------+---------------------------------------------------------------------------------+
| Rel Hum               | This is the relative humidity given to the nearest whole percentage             |
+-----------------------+---------------------------------------------------------------------------------+

Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.

+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| Week No. & Overview                                                                 | Individual Weekly Tasks                                                    | Assigned Team Members       |
+:===================================================================================:+:==========================================================================:+:===========================:+
| [**Week 1\                                                                          | Setting up environment, data blending                                      | *Dhanya, Udit*              |
| **]{.underline}11/8/23 - 11/15/23                                                   |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Objective:**]{.underline}\                                                       | Identify missing values and perform imputation to deal with missing values | *Riyanshi, Ajinkya*         |
| Data cleaning & data pre-processing                                                 |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Deliverables:**]{.underline}\                                                    | Data type conversions                                                      | *Kendall*                   |
| Cleaned dataset ready for exploratory analysis.                                     |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| \-                                                                                  | \-                                                                         | \-                          |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Week 2\                                                                          | Visualization of key metrics                                               | *Udit, Dhanya*              |
| **]{.underline}11/15/23 - 11/22/23                                                  |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Objective:**]{.underline}\                                                       | Feature engineering and selection                                          | *Riyanshi*                  |
| Gain insights into the dataset and prepare for model development.                   |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Deliverables:**]{.underline}\                                                    | Data normalization and standardization                                     | *Kendall, Ajinkya*          |
| A thorough analysis of data insights and prepared features for modeling.            |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| \-                                                                                  | \-                                                                         | \-                          |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Week 3\                                                                          | Implement anomaly detection algorithms                                     | *Riyanshi, Kendall, Dhanya* |
| **]{.underline}11/22/23 - 11/29/23                                                  |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Objective:**]{.underline}\                                                       | Validate and analyze anomalies                                             | *Udit, Ajinkya*             |
| Identify and address anomalies in the dataset and draw conclusions concerning UHIs. |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Deliverables:**]{.underline}\                                                    | \-                                                                         | \-                          |
| Anomaly detection results and cleaned dataset.                                      |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| \-                                                                                  | \-                                                                         | \-                          |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Week 4\                                                                          | Build classification models                                                | *Dhanya, Kendall*           |
| **]{.underline}11/29/23 - 12/6/23                                                   |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Objective:**]{.underline}\                                                       | Model optimization & cross-validate models                                 | *Riyanshi, Udit*            |
| Develop classification models for anomalies and evaluate their performance.         |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Deliverables:**]{.underline}\                                                    | Evaluation metrics and performance analysis                                | *Ajinkya*                   |
| Working models and performance evaluation report.                                   |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| \-                                                                                  | \-                                                                         | \-                          |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Week 5\                                                                          | -                                                                          | \-                          |
| **]{.underline}12/6/23 (Presentation)                                               |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Objective:**]{.underline}\                                                       | Compile project documentation                                              | *Kendall, Dhanya*           |
| Document the project process and results, and prepare a presentation.               |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+
| [**Deliverables:**]{.underline}\                                                    | Prepare final presentation                                                 | *Udit, Riyanshi, Ajinkya*   |
| Complete project documentation and ready-to-deliver presentation.                   |                                                                            |                             |
+-------------------------------------------------------------------------------------+----------------------------------------------------------------------------+-----------------------------+

The primary objective of this project is to leverage data mining and machine learning algorithms to detect and characterize Urban Heat Islands (UHIs) across three different areas in the Dallas region. By analyzing weather pattern anomalies and temperature distributions, the project aims to gain a deeper understanding of the micro-climatic effects of urbanization.

-   such as prolonging heat waves and contributing to heavy precipitation. This effect occurs when natural land and vegitation is replaced with pavement, buildings, and various surfaces that absorb and retain heat. So, high-rise buildings act as heat sinks during the day and release heat after sunset, prolonging the increase of temperatures.

UHI is a term used in the field of climate science, "climatology," to describe the subtle yet powerful effect that a city can have on amplifying or suppressing weather conditions, such as prolonging heat waves and heavy precipitation. A UHI is formed when natural land and vegetation are replaced by buildings, pavements, and various surfaces that absorb and retain heat. So, for example, a high-rise building will act as a heat sink during the day and then release that energy after sunset, prolonging the heat and humidity conditions which less dense and suburban areas do not have the problem.

UHI is the result of dense buildings and pavements covering or replacing natural land and vegetation are replaced by buildings, pavements, and various surfaces that absorb and retain heat. A high-rise building will act as a heat sink during the day, then release that energy after sunset--a problem that suburban areas do not face.

-   when cities replace natural land cover with dense concentrations of pavement, buildings, and other surfaces that absorb and retain heat

-   High-rise buildings act as heat sinks during the day and release heat after sunset, so temperatures actually increase at night for UHI. (Auburn University)

-   We will be studying the daily climatological data to classify the level of intensity of UHI for our region of study, as well as to do any trend analysis on the climate data to spot any trends on urban vs. rural temperature, etc.

-   Urban Heat Islands is a term to describe urban areas that are abundant with buildings but less abudant with vegetation, which helps with regulating humidity levels. In our case, choosing a city like Dallas, with a high humidity, will see a greater heat intensity than a less-UHI.

-   because there is less vegetation in urban areas,

<https://www.metlink.org/fieldwork-resource/urban-heat-island-introduction/>

The primary objective of this project is to study and analyze climatological data for three cities inside and outside of the Dallas, TX region, and to classify these cities based on an "Urban Heat Island" (UHI) scale.

UHI is a term used in the field of climate science, "climatology," to describe the subtle yet powerful effect that a city can have on amplifying or suppressing weather conditions---heat waves, precipitation, etc. A UHI is formed when natural land and vegetation are replaced by buildings, pavements, and surfaces. These objects, particularly high-rise buildings, will absorb heat throughout the day then release the energy after sunset, thus prolonging hot and humid conditions. This is known as a "micro climate" in climatology since a UHI is unintentionally affecting atmospheric conditions through its own, dense infrastructure---creating its own micro climate.

By analyzing the micro climate data for Dallas, a major urban city in the country, and two suburban cities outside of Dallas, we can create a classification system for different types of UHIs and label these cities accordingly. Additionally, we can perform trend analysis by comparing and contrasting various weather patterns between different types of UHIs.

```{r}

#| include: false

# Question 1, Part 1: k-Means Clustering, SVM (Support Vector Machines), Isolation Forest and Local Outlier Factor (LOF), Z-scores. First 3 are models.

# Question 1, Part 2: Time-series analysis (data viz & dashboards; heat maps).
```

UHI, a term used in the field of climate science, occurs when a city replaces land and natural vegetation with buildings, pavements and surfaces that absorbs and holds heat. This effect is also known as "micro climate" since a city's infrastructure is creating its own local weather by amplifying or suppressing conditions such as heat waves, precipitation, etc.
