Make sure to load the data and use inline code for some of this information.

Byte Busters' Draft:

The Byte Busters will be using data from the National Centers for Environmental Information

(https://www.ncei.noaa.gov/)...

Udit & Ajinkya:

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

[Riyanshi]{.underline}:

-   *Leverage advanced data mining techniques to identify and analyze urban heat islands and weather anomalies, providing insights into the impacts of urbanization on local climates.*

1.  Week 1: Data Cleaning and Data Preprocessing (11/8/2023 - 11/15/2023)

2.  Week 2: Exploratory Data Analysis (11/15/2023 - 11/22/2023)

3.  Week 3: Anomaly Detection (11/22/2023 - 11/29/2023)

4.  Week 4: Model Development and Evaluation (11/29/2023 - 12/6/2023)

5.  Week 5: Documentation & Presentation (12/6/2023 - 13/6/2023)

\-\-\-\--

This is what Anjik sent to me on WhatsApp:

data1_rows, data1_cols = dallas_data.shape\
data2_rows, data2_cols = arlington_data.shape\
data3_rows, data3_cols = denton_data.shape

```{r}
#Dallas Dataset: 

dim(dallas_data)

dallas_dimension <- data.frame(dallas_data)
dallas_dimension
```

Anjinka & Udit

wind conditions, precipitation levels, degree days, and temperature extremes. Additional data is included in the tabulated monthly summaries: maximum, minimum, and average temperatures; deviations from normal; dew point temperatures; average station pressure; ceiling height; visibility; weather patterns; wet bulb temperatures; relative humidity; heating and cooling degree days; daily precipitation; average wind speeds; fastest wind speeds and their directions; sky cover; and instances of sunshine, snowfall, and snow depth.

1.  What methods can be employed to assess variations in the intensity of urban heat islands across diverse metropolitan and rural regions, and how can we characterize the temporal patterns of these variations throughout a specific year?

<!-- -->

1.  How can we pinpoint distinct anomalies that are typical of urban heat islands, and what is the extent of their correlation with urban infrastructure and population density?

#### Dallas 2022 Dataset

`{r} #| label: dallas-dataset #| echo: false #| warning: false  dallas_data <- read_csv(here('./data/dallas-2022.csv'))  # Dimensions of dataset: num_rows <- nrow(dallas_data) num_columns <- ncol(dallas_data)  cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")   dallas_data |>   head() |>   formattable()}`

#### **Arlington 2022 Dataset**

`{r} #| label: arlington-dataset #| echo: false #| warning: false  arlington_data <- read_csv(here('./data/arlington-2022.csv'))  # Dimensions of dataset: num_rows <- nrow(arlington_data) num_columns <- ncol(arlington_data)  cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")  arlington_data |> # Need to format "DATE" and other columns if we have time.   head() |>   formattable()}`

### **Denton 2022 Dataset**

```{r}
#| label: denton-dataset
#| echo: false
#| warning: false

denton_data <- read_csv(here('./data/denton-2022.csv'))

# Dimensions of dataset:
num_rows <- nrow(denton_data)
num_columns <- ncol(denton_data)

cat("(", num_rows, " rows x ", num_columns, " columns)", sep = "")

denton_data |> # Need to format "DATE" and other columns if we have time.
  head() |>
  formattable()
```

-   **data/:** Used for storing any necessary data files for the project, such as input files.

-   **images/:** Used for storing plots of our project.

-   **presentation_files/:** Folder for having presentation related files.

-   **\_extra/:** Used to brainstorm our analysis which won't impact our project workflow.

-   **\_freeze/:** This folder is used to store the generated files during the build process. These files represent the frozen state of the website at a specific point in time.

-   **\_site/:** Folder used to store the generated static website files after the site generator processes the quarto document.

-   **.github/:** Folder for storing github templates and workflow.

| Columns               | Description                                                                     |
|----------------------------------------------|--------------------------|
| Wind Gust Speed       | Wind gusts occurring during time of observation. Given in miles per hour (mph)  |
| Dry Bulb Temperature  | The measure of air temperature (Fahrenheit)                                     |
| Precipitation Levels  | Total Liquid Content (TLC) Water equivalent amount of precipitation for the day |
| Dew Point Temperature | (Add description)                                                               |
| Pressure - Avg SL     | Daily average sea level pressure                                                |
| Rel Hum               | This is the relative humidity given to the nearest whole percentage             |
